Graph Generators: State of the Art and Open Challenges

### 2019-08-06 challenges.tex
Irena's pass (plagiarism) - th rest of the text

The evolution of graphs can be considered from multiple perspectives. We can
assume a static set of nodes and a varying set of edges.  Or, there are
applications where the graphs only ``grow'', i.e., the set of nodes and/or edges
is only extended with new items. In the most general case we can assume any
changes in both set of nodes and set of edges. Anyway with the evolution aspect
the complexity of classical graph problems increases
significantly~\cite{Michail2015,Wu:2014:PPT:2732939.2732945}. In some graph
applications, such as, e.g., social networks, the evolution of the data is a
significant aspect, especially in the activity
graphs~\cite{doreian1997evolution,Kumar:2006:SEO:1150402.1150476,Hellmann2014583,wang2013,Kossinets88,Viswanath:2009:EUI:1592665.1592675}.
However, as shown
in~\cite{Leskovec:2005:RMT:2101235.2101254,Leskovec:2005:GOT:1081870.1081893},
evolving graphs have further specific features. For example, some graphs grow
over time according to a \emph{densification power law} which means that in real
graphs, edges tend to appear at a higher pace than vertices, meaning that these
graphs densify as they grow. Also the way the new edges are distributed has the
effect of a shrinking diameter that ends up stabilizing as the graph grows with time.

@@

The respective data generator should hence be able to simulate a natural growth
and/or changes in the structure of the graph with regards to the various
features of distinct use cases. However,  even though the area of dynamic graphs
is intensively studied, surprisingly there seem to exist only very few proposals
of a generator for dealing with this area.
In~\cite{GoerkeKlugeSchumm2012_1000029825} the authors focus on \emph{clustering
dynamic graphs}, i.e. graphs where the clustering corresponds to the partition
of nodes into natural groups based on the concept of intra-cluster
versus inter-cluster density of edges. The generator generates a time series
of random graphs $G_0, G_1, ..., G_n$, where $G_t$ emerges from $G_{t-1}$ via
successive atomic updates like for instance, the insertion of a vertex or the removal of an
edge. The generator dynamically monitors the ground truth clustering, and the probability
of the updates is chosen in such a way that the ground truth is
maintained while the randomness of the generated graph is kept.

@@

\subsection{Multi-Model Data}
With the dawn of Big Data and especially its Variety, another key 3V characteristic, new types of
database management systems have emerged. One of the most interesting ones
are the so-called \emph{multi-model databases}~\cite{Lu:2019:MDN:3341324.3323214} that enable to store and
thus query across structurally different data, including structured, semi-structured, and unstructured. There exist various types of
multi-model systems combining  distinct subsets of Big Data structures including graph data.
For example, OrientDB\footnote{\url{http://orientdb.com/orientdb/}} which has been mainly designed as an object DBMS currently supports graph, document,
key/value, and object models. Such type of DBMSs also needs a specific
data generator that would enable
to test new features and analyze efficiency of operations. However, since the
multi-model systems are in the context of Big Data rather new, there exist only
a few benchmarks targeting multi-model DBMSs (such as
Bigframe~\cite{journals/pvldb/KunjirKB14} or UniBench~\cite{conf/cidr/lu17})
with limited capabilities.

Another interesting approach to multi-model data is to adopt a unifying
expressive graph data model, known as property graph data
model~\cite{BFVY18}. Such a model allows to specify multi-edges and list of
properties for the nodes. Synthetic graph generators for property graphs and its
companion standard graph query language~\cite{Angles18,AnglesABBFGLPPS18} are also needed in order to boost
their availability and adoption for different communities.

@@

With the advent of neural networks and specially generative adversarial networks
(GANs)~\cite{goodfellow2014generative}, several researchers have started to
explore their application to generate graphs. This is the case of
~\cite{kipf2016variational,grover2018graphite,simonovsky2018graphvae,li2018learning,you2018graphrnn},
which present several generative models to generate realistic graphs.  Such
techniques still suffer from several problems. For instance, some of them are
limited to learn from a single
graph~\cite{kipf2016variational,grover2018graphite} or generate small
graphs~\cite{simonovsky2018graphvae,li2018learning,you2018graphrnn}. The
technique proposed in~\cite{you2018graphrnn} is capable of generating graphs
with complex edge dependencies (e.g. community structure) and is not restricted
to graphs of a fixed size. However, there are still in general several open
challenges, including the capability of learning from and generating large graphs comparable
in size to those typically used for benchmarking, and robust generation
techniques with structural guarantees (e.g. degree distribution, clustering
coefficient, etc.).

>>>>

The evolution of graphs can be considered from multiple perspectives. We can
assume a static set of nodes and a varying set of edges.  Or, there are
applications where the graphs only ``grow'', i.e., the set of nodes and/or edges
is only extended with new items. In the most general case we can assume any
changes in both set of nodes and set of edges. Anyway with the evolution aspect
the complexity of classical graph problems increases
significantly~\cite{Michail2015,Wu:2014:PPT:2732939.2732945}. In some graph
applications, such as, e.g., social networks, the evolution of the data is a
significant aspect, especially in the activity
graphs~\cite{Kumar:2006:SEO:1150402.1150476,doreian1997evolution,Hellmann2014583,wang2013,Viswanath:2009:EUI:1592665.1592675,Kossinets88}.
However, as shown
in~\cite{Leskovec:2005:RMT:2101235.2101254,Leskovec:2005:GOT:1081870.1081893},
evolving graphs have further specific features. For example, some graphs grow
over time according to a \emph{densification power law} which means that in real
graphs, edges tend to appear at a higher pace than vertices, meaning that these
graphs densify as they grow. Also the way the new edges are distributed has the
effect of a shrinking diameter that ends up stabilizing as the graph grows with time.

@@

The respective data generator should hence be able to simulate a natural growth
and/or changes in the structure of the graph with regards to the various
features of distinct use cases. However,  even though the area of dynamic graphs
is intensively studied, surprisingly there seem to exist only very few proposals
of a generator for dealing with this area.
In~\cite{GoerkeKlugeSchumm2012_1000029825} the authors focus on \emph{clustering
dynamic graphs}, i.e. graphs where the clustering corresponds to the partition
of nodes into natural groups based on the concept of density of edges within and between the clusters. The generator generates a time series
of random graphs $G_0, G_1, ..., G_n$, where $G_t$ emerges from $G_{t-1}$ via
successive atomic updates like for instance, the insertion of a vertex or the removal of an
edge. The generator dynamically monitors the ground truth clustering, and the probability
of the updates is chosen in such a way that the ground truth is
maintained while the randomness of the generated graph is kept.

@@

\subsection{Multi-Model Data}
With the dawn of Big Data and especially its Variety, another key 3V characteristic, new types of
database management systems have emerged. One of the most interesting ones
are the so-called \emph{multi-model databases}~\cite{Lu:2019:MDN:3341324.3323214} that enable to store and
thus query across structurally different data, including unstructured, semi-structured, and structured. There exist various types of
multi-model systems combining  distinct subsets of Big Data structures including graph data.
For example, OrientDB\footnote{\url{http://orientdb.com/orientdb/}} which has been mainly designed as an object DBMS currently supports graph, document,
key/value, and object models. Such type of DBMSs also needs a specific
data generator that would enable
to test new features and analyze efficiency of operations. However, since the
multi-model systems are in the context of Big Data rather new, there exist only
a few benchmarks targeting multi-model DBMSs (such as
Bigframe~\cite{journals/pvldb/KunjirKB14} or UniBench~\cite{conf/cidr/lu17})
with limited capabilities.

Another interesting approach to multi-model data is to adopt a unifying
expressive graph data model, so-called \emph{property graph data
model}~\cite{BFVY18}. Such a  model allows to specify multi-edges and list of
properties for the nodes. Synthetic graph generators for property graphs and its
companion standard graph query language~\cite{Angles18,AnglesABBFGLPPS18} are also needed in order to boost
their availability and adoption for different communities.

@@

With the advent of neural networks and specially generative adversarial networks
(GANs)~\cite{goodfellow2014generative}, several researchers have started to
explore their application to generate graphs. This is the case of
~\cite{simonovsky2018graphvae,kipf2016variational,grover2018graphite,li2018learning,you2018graphrnn},
which present several generative models to generate realistic graphs.  Such
techniques still suffer from several problems. For instance, some of them are
limited to learn from a single
graph~\cite{kipf2016variational,grover2018graphite} or generate small
graphs~\cite{simonovsky2018graphvae,you2018graphrnn,li2018learning}. The
technique proposed in~\cite{you2018graphrnn} is capable of generating graphs
with complex edge dependencies (e.g. community structure) and is not restricted
to graphs of a fixed size. However, there are still in general several open
challenges, including the capability of learning from and generating large graphs comparable
in size to those typically used for benchmarking, and robust generation
techniques with structural guarantees (e.g. degree distribution, clustering
coefficient, etc.).

### 2019-08-06 generators_SocialNetworks.tex
Irena's pass (plagiarism) - th rest of the text

\paragraph{S3G2} The Scalable Structure-correlated Social Graph Generator
(S3G2)~\cite{Pham2013} is a general framework which produces a directed labeled
graph whose vertices represent objects having property values. The respective
classes determine the structure of the properties. S3G2 does not aim at
generating near real-world data, but at generating synthetic graphs with a
correlated structure. Hence, the existing data values influence the probability
of choosing a certain property value from a pre-defined dictionary, or the
probability of connecting two nodes. For example, the degree distribution can be
correlated with the properties of a node and thus, e.g., people having many
friend relationships typically post more comments and pictures. The data
generation process starts with generating a number of nodes with property values
generated according to specified property value correlations and then adding
respective edges according to specified correlation dimensions. It has multiple
phases, each focusing on one correlation dimension. Data is generated in a Map
phase corresponding to a pass along one correlation dimension. Then the data are
sorted along the correlation dimension in the following Reduce phase. A
heuristic observation that ``the probability that two nodes are connected is
typically skewed with respect to some similarity between the nodes'' enables to
focus only on sliding window of most probable candidates. The core idea of the
framework is demonstrated using an example of a social network (consisting of
persons and social activities).  The dictionaries for property values are
inspired by DBpedia and provided with 20 property value correlations. The edges
are generated according to 3 correlation dimensions.

\paragraph{SIB} The developers of the Social Network Intelligence BenchMark (SIB)\footnote{\url{https://www.w3.org/wiki/Social_Network_Intelligence_BenchMark}} based the design of their benchmark on the claim that existing benchmarks are limited in reflecting the characteristics of the real RDF databases and are mostly focusing on the relational style aspects. Hence, they proposed a benchmark for  query evaluation using real graphs~\cite{sib}. The proposed benchmark mimics using an RDF store for a social network. \iffalse site where users and their interactions form a social graph of social activities such as creating/managing groups, writing posts and posting comments.\fi The distribution of the generated data for each type follows the  distribution of the associated type inferred from real-world social networks. Additionally, association rules are exploited for representing the real-world data correlation in the generated synthetic data. The  generated data is linked with the RDF datasets from DBpedia. The benchmark specification contains 3 query mixes -- interactive, update, and analysis -- expressed in SPARQL 1.1 Working Draft.

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduces two synthetic generators to reproduce two characteristics typically
observed in social networks: node features and multiple link types. Both
generators extend  the generator proposed by~\cite{wang2011leveraging}.
which starts with a small number of
nodes and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes with the
same label; these labels can be considered as being equivalent to community
membership.

@@

\paragraph{LDBC SNB} Despite having a common Facebook-like dataset, thanks to
three distinct workloads the Social Network Benchmark
(SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC represents three
distinct benchmarks. The network nodes correspond to people and the edges
represent their friendship and messages they post in discussion trees on their
forums. The three query workloads involve: (1) SNB-Interactive, i.e., complex
read-only queries accessing a significant portion of data, (2) SNB-BI, i.e.,
queries accessing a high percentage of  entities and grouping them in various
dimensions, and (3) SNB-Algorithms, involving graph analysis algorithms, such as
community detection, PageRank, BFS, and clustering. The graph generator, called
Datagen, is a fork of S3G2 ~\cite{Pham2013} and realizes power laws, uses skewed
value distributions, and ensures reasonable correlations between graph
structures and property values. Additionally, it extends S3G2 with "spiky"
patterns in the distribution of social network activity along the timeline, also
provides the ability of generating update streams to the social network. Datagen
is also based on Hadoop in order to provide scalability, but comparet to S3G2 it
contains numerous performance improvements and the ability to be deterministic
regardless of the number of computer nodes used for the generation of the graphs
and for a given set of configuration parameters.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the majority of existing works focuses on
topology generation which approximates the characteristics of a real-world social
network (e.g., a small graph diameter, skew degree
distribution, small average path length, and community structures); however,  this is usually done without any data. Hence, they
introduced a general stochastic modeling approach that enables the users to
populate a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
that describes distribution profiles, attribute value definitions, using a
parameterizable set of data propagation rules and affinities, and (3) data
population.

>>>>

\paragraph{S3G2} The Scalable Structure-correlated Social Graph Generator
(S3G2)~\cite{Pham2013} is a general framework which produces a directed labeled
graph whose vertices represent objects having property values. The respective
classes determine the structure of the properties. S3G2 does not aim at
generating near real-world data, but at generating synthetic graphs with a
correlated structure. Hence, the existing data values influence the probability
of choosing a particular property value from a pre-defined dictionary or connecting two nodes. For example, the degree distribution can be
correlated with the properties of a node and thus, e.g., people who have many
friend relationships typically post more comments and pictures. The data
generation process starts with generating a number of nodes with property values
generated according to specified property value correlations and then adding
respective edges according to specified correlation dimensions. It has multiple
phases, each focusing on one correlation dimension. Data is generated in a Map
phase corresponding to a pass along one correlation dimension. Then the data are
sorted along the correlation dimension in the following Reduce phase. A
heuristic observation that ``the probability that two nodes are connected is
typically skewed with respect to some similarity between the nodes'' enables to
focus only on sliding window of most probable candidates. The core idea of the
framework is demonstrated using an example of a social network (consisting of
persons and social activities).  The dictionaries for property values are
inspired by DBpedia and provided with 20 property value correlations. The edges
are generated according to 3 correlation dimensions.

\paragraph{SIB} The developers of the Social Network Intelligence BenchMark (SIB)\footnote{\url{https://www.w3.org/wiki/Social_Network_Intelligence_BenchMark}} based the design of their benchmark on the claim that the state-of-the-art benchmarks are limited in reflecting the characteristics of the real RDF databases and are mostly focusing on the relational style aspects. Hence, they proposed a benchmark for  query evaluation using real graphs~\cite{sib}. The proposed benchmark mimics using an RDF store for a social network. \iffalse site where users and their interactions form a social graph of social activities such as creating/managing groups, writing posts and posting comments.\fi The distribution of the generated data for each type follows the  distribution of the associated type inferred from real-world social networks. Additionally, association rules are exploited for representing the real-world data correlation in the generated synthetic data. The  generated data is linked with the RDF datasets from DBpedia. The benchmark specification contains 3 query mixes -- interactive, update, and analysis -- expressed in SPARQL 1.1 Working Draft.

\paragraph{Cloning of Social Networks} ~\cite{Sukthankar-SocialInfo2014}
introduces two synthetic generators to reproduce two characteristics typically
observed in social networks: node features and multiple link types. Both
generators extend  the generator proposed by~\cite{wang2011leveraging}.
which starts with a small number of
nodes and new nodes are added until the network reaches the required number. It
has two basic parameters: homophily and link density. A high \emph{homophily}
value reflects that links have higher chances to be established among the nodes belonging to the same community, 
whereas the community membership is represented by the same labels.

@@

\paragraph{LDBC SNB} Despite having a common Facebook-like dataset, thanks to
three distinct workloads the Social Network Benchmark
(SNB)~\cite{Erling:2015:LSN:2723372.2742786} provided by LDBC represents three
distinct benchmarks. The network nodes correspond to people and the edges
represent their friendship and messages they post in discussion trees on their
forums. The three query workloads involve: (1) SNB-Interactive, i.e., complex
read-only queries accessing a high portion of data, (2) SNB-BI, i.e.,
queries accessing a high percentage of  entities and grouping them in various
dimensions, and (3) SNB-Algorithms, involving graph analysis algorithms, such as
community detection, PageRank, BFS, and clustering. The graph generator, called
Datagen, is a fork of S3G2 ~\cite{Pham2013} and realizes power laws, uses skewed
value distributions, and ensures reasonable correlations between graph
structures and property values. Additionally, it extends S3G2 with "spiky"
patterns in the distribution of social network activity along the timeline, also
provides the ability of generating update streams to the social network. Datagen
is also based on Hadoop in order to provide scalability, but comparet to S3G2 it
contains numerous performance improvements and the ability to be deterministic
regardless of the number of computer nodes used for the generation of the graphs
and for a given set of configuration parameters.

@@

\paragraph{Towards More Realistic Data} \cite{Nettleton2016} argued that the majority of existing works focuses on
topology generation which approximates the features of a real-world social
network (e.g., community structures, skew degree
distribution, a small average path length, or a small graph diameter); however,  this is usually done without any data. Hence, they
introduced a general stochastic modeling approach that enables the users to
fill a graph topology with data. The approach has three steps: (1) topology
generation (using R-MAT) plus community identification using the Louvain
method~\cite{1742-5468-2008-10-P10008} or usage of a real-world topology from
SNAP\footnote{\url{https://snap.stanford.edu/data/}}, (2) data definition
that describes  definitions of attribute values (distribution profiles) using a
parameterizable set of affinities and data propagation rules, and (3) data
population.

### 2019-08-06 generators_community_detection.tex
Irena's pass (plagiarism) - th rest of the text

Community detection is one of the many graph analytics algorithms typically used
in domains such as social networks or bioinformatics. \emph{Communities} are groups of nodes that are highly mutually connected, while being scarcely connected to
the rest of the graph. Such communities emerge from the fact that real-world graphs
are not random, but follow real-world dynamics that make similar entities to
have a larger probability to be connected. As a consequence, detected
communities are used to reveal vertices with similar characteristics, for
instance to discover functionally equivalent proteins in protein-to-protein
interaction networks, or persons with similar interests in social networks. Such
applications have made community detection a hot topic during the last 15
years with tens of developed algorithms and detection
strategies~\cite{doi:10.1002/wics.1403,Kim:2015:CDM:2854006.2854013}. For
comparing the quality of the different proposed techniques, one needs graphs
with \emph{reference communities}, that is, communities known beforehand. Since
it is very difficult to have large real-world graphs with reference communities
(mainly because these would require a manual labeling), graphs for benchmarking
community detection algorithms are typically generated synthetically.

@@

\paragraph{LFR} Lancichinetti, Fortunato
and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark
graphs for community detection where communities are of diverse sizes and
densities. The generated communities follow a power-law distribution whose parameters can be configured. The degree of the
nodes is also sampled from a power-law distribution. Additionally, the generator
introduces the concept of the ``mixing factor'', which consists of the percentage of
edges in the graph connecting nodes that belong to different communities. Such parameter
allows  the degree of modularity of the generated graph  to be tuned, thus
testing the robustness of the algorithms under different conditions. The
generation process is implemented as an optimization process starting with an empty graph and
progressively filling it with nodes and edges guided by the specified constraints.

@@

\paragraph{Stochastic Block Models} Another popular family of generators
widely used in the community detection field are the stochastic block
models~\cite{holland1983stochastic}. In such models, the community structure of
the graph is typically defined as an array of $n$ community or cluster sizes
and a density square matrix of size $n \times n$ containing the density of
intra-cluster edges (in the diagonal of the matrix) and the density of
inter-cluster edges. Then, a stochastic procedure is run to sample graphs
from such array and matrix, using the sizes to compute the possible edges
and the densities as probabilities of such edges to exist. The popularity of
these methods stem from its simplicity and scalability, which makes them
suitable for generating large graphs fast and in a distributed environments,
provided that the density matrix is sparse (as it happens in most of real-world
                                            world graphs). Moreover, given
that the generation process of such models is mathematically tractable, they
are typically used to analyze the limitations of community detection
algorithms such as those based on modularity
optimization~\cite{fortunato2007resolution} or based on
triads~\cite{prat2016put}. Extensions of such models exist, such as the
Mixed Membership Stochastic Block Model~\cite{airoldi2008mixed}, for
overlapping communities.

>>>>

Community detection is one of the many graph analytics algorithms typically used
in domains such as social networks or bioinformatics. \emph{Communities} are usually defined as sets of nodes that are highly mutually connected, while being scarcely connected to
the other nodes of the graph. Such communities emerge from the fact that real-world graphs
are not random, but follow real-world dynamics that make similar entities to
have a larger probability to be connected. As a consequence, detected
communities are used to reveal vertices with similar characteristics, for
instance to discover functionally equivalent proteins in protein-to-protein
interaction networks, or persons with similar interests in social networks. Such
applications have made community detection a hot topic during the last 15
years with tens of developed algorithms and detection
strategies~\cite{doi:10.1002/wics.1403,Kim:2015:CDM:2854006.2854013}. For
comparing the quality of the different proposed techniques, one needs graphs
with \emph{reference communities}, that is, communities known beforehand. Since
it is very difficult to have large real-world graphs with reference communities
(mainly because these would require a manual labeling), graphs for benchmarking
community detection algorithms are typically generated synthetically.

@@

\paragraph{LFR} Lancichinetti, Fortunato
and Radicchi (hence LFR)~\cite{PhysRevE.78.046110} propose a class of benchmark
graphs for community detection where communities are of diverse sizes and
densities. The generated communities follow a power-law distribution whose parameters can be configured. The degree of the
nodes is also sampled from a power-law distribution. Additionally, the generator
introduces the concept of the ``mixing factor'', which consists of the percentage of
edges in the graph connecting nodes that belong to distinct communities. Such parameter
allows  the degree of modularity of the generated graph  to be tuned, thus
testing the robustness of the algorithms under different conditions. The
generation process is implemented as an optimization process starting with an empty graph and
progressively filling it with nodes and edges guided by the specified constraints.

@@

\paragraph{Stochastic Block Models} Another popular family of generators
widely used in the community detection field are the stochastic block
models~\cite{holland1983stochastic}. In such models, the community structure of
the graph is typically defined as an array of $n$ community or cluster sizes
and a density square matrix of size $n \times n$ containing the density of
intra-cluster edges (in the diagonal of the matrix) and the density of
inter-cluster edges. Then, a stochastic procedure is run to sample graphs
from such array and matrix, using the sizes to compute the possible edges
and the densities as probabilities of such edges to exist. The popularity of
these methods stem from its simplicity and scalability, which makes them
suitable for generating large graphs fast and in a distributed environments,
provided that the density matrix is sparse (as it happens in most of real-world
                                            world graphs). Moreover, given
that the generation process of such models is mathematically tractable, they
are typically used to analyze the limitations of algorithms for community detection
 such as those based on modularity
optimization~\cite{fortunato2007resolution} or based on
triads~\cite{prat2016put}. Extensions of such models exist, such as the
Mixed Membership Stochastic Block Model~\cite{airoldi2008mixed}, for
overlapping communities.
